# Welcome {-}

```{r "intro"}
#| echo: false
.gh_url <- file.path('https://github.com', rmarkdown::metadata[['github-repo']])
pkglines <- read.dcf("DESCRIPTION")
pkg <- pkglines[, "Package"][[1]]
version <- pkglines[, "Version"][[1]]
authors <- eval(parse(text = pkglines[, "Authors@R"]))
license <- pkglines[, "License"][[1]]
intro <- glue::glue(
    "**Package:** {pkg}<br/>\n", 
    "**Authors:** {paste(format(authors, include = c('given', 'family', 'role')), collapse = ', ')}<br/>\n", 
    "**Compiled:** {as.character(Sys.Date())}<br/>\n", 
    "**Package version:** {version}<br/>\n", 
    "**R version:** <b style='color:red;'>{R.version.string}</b><br/>\n", 
    "**BioC version:** <b style='color:red;'>{BiocManager::version()}</b><br/>\n", 
    "**License:** {license}<br/>"
)
```

`r intro`

This is the landing page for the **"Epigenomics Data Analysis"** 
workshop, ed. 2023.  

## What {-}

This course will introduce biologists and bioinformaticians to the field of 
regulatory epigenomics. We will cover a range of software and analysis 
workflows for processing of next-generation sequencing datasets and 
quantitative analysis ChIP-seq, ATAC-seq data and RNA-seq data. Towards the end of the workshop, 
a brief introduction to chromatin conformation capture (Hi-C) experiments and analysis 
will emphasize multi-omics data integration.

We will start by introducing general concepts about epigenomics. 
From there, we will then continue to describe the main analysis steps to go 
from raw sequencing data to processed and usable data. We will present 
classical analysis workflows, their output and the possible paths to 
investigate downstream of this.  

Throughout the workshop, `bash` tools and `R/Bioconductor` packages will 
be used to analyse datasets and learn new approaches. 

## Who {-}

The course is aimed at researchers interested in learning how to extract 
biological insights from genomics data, such as ChIP-seq, ATAC-seq or Hi-C.  

It is primarily targeting researchers who are new to the field 
of bioinformatics, with practical experience in the experimental side of 
epigenomics. 

Attendees should have a background in biology as well as be somewhat familiar with 
genomic data and common file formats from NGS sequencing experiments 
(fastq, BAM, BED).

Practical exercises will use command-line Linux and R code and will be presented 
as notebooks to ensure reproducible coding.

## Why {-}

At the end of this course, you should be able to:

- Understanding important genomic file formats
- Process most types of genomic datasets (RNA-seq, ATAC-seq, Hi-C, ChIP-seq, ...)
- Analyzing processed datasets to extract relevant information and answer biological questions
- Good practices to avoid confounding variables and pitfalls in the processing.
- Proper use of controls and normalization.
- Integration of different sequencing "omics" datasets
- Characterisation of the global 3D structures from the sequencing data
- Detection of regulatory interactions and quantification of their changes between conditions.

Throughout the course, we will also have a focus on reproducible research, documented content and interactive reports.

## Instructors {-}

[Jacques Serizay](mailto:jacques.serizay@pasteur.fr)

# Docker image {-}

A `Docker` image built from this repository is available here: 

👉 [ghcr.io/js2264/biocbook.epigenomics](https://ghcr.io/js2264/biocbook.epigenomics) 🐳

::: {.callout-tip icon='true'}

## Get started now 🎉

You can get access to all the packages used in this book in < 1 minute, 
using this command in a terminal: 

```{sh "docker", filename="bash"}
#| eval: false
docker run -it ghcr.io/js2264/biocbook.epigenomics:devel R
```

:::

# RStudio Server {-}

An RStudio Server instance can be initiated from the `Docker` image as follows: 

```{sh "rstudio", filename="bash"}
#| eval: false
docker run \
    --volume <local_folder>:<destination_folder> \
    -e PASSWORD=OHCA \
    -p 8787:8787 \
    ghcr.io/js2264/biocbook.epigenomics:devel
```

The initiated RStudio Server instance will be available at 
[https://localhost:8787](https://localhost:8787).

# Session info {-}

::: {.callout-note collapse="true"}

## Click to expand 👇

```{r "session info"}
#| cache: false
sessioninfo::session_info(
    installed.packages()[,"Package"], 
    include_base = TRUE
)
```

:::
